Light versus dark in strong-lens galaxies: Dark matter haloes that are rounder than their stars

### 2015-01-30 shapes/paper.tex
Paper submitted

% \def\ltsima{$\; \buildrel < \over \sim \;$}
% \def\simlt{\lower.5ex\hbox{\ltsima}}
% \def\gtsima{$\; \buildrel > \over \sim \;$}
% \def\simgt{\lower.5ex\hbox{\gtsima}}
\def\Glass{{\sc Glass}}
\def\PixeLens{{\sc PixeLens}}
\newcommand{\cb}[1]{{\color{red} \textbf{#1}}}

@@

\author[Bruderer et al.]
{\parbox{\textwidth}{Claudio Bruderer,$^{1}$\thanks{E-mail: \texttt{claudio.bruderer@phys.ethz.ch}}
Justin I. Read,$^{1,2}$
Jonathan Coles,$^{3}$
\cb{Dominik Leier,$^{4}$
Emilio E. Falco,$^{5}$
Ignacio Ferreras$^{6}$} and
Prasenjit Saha,$^{7}$}\vspace{0.4cm}\\
\parbox{\textwidth}{$^{1}$Institute for Astronomy, Department of Physics, ETH Zurich, Wolfgang-Pauli-Strasse 27, 8093 Z\"urich, Switzerland\\
$^{2}$Department of Physics, University of Surrey, Guildford, GU2 7XH, UK\\
$^{3}$Exascale Research Computing Lab, Campus Teratec, 2 Rue de la Piquetterie, 91680 Bruyeres-le-Chatel, France\\
$^{4}$Dipartimento di Fisica e Astronomia, Universit\`{a} di Bologna, viale Berti Pichat 6/2, 40127, Bologna, Italy\\
$^{5}$Harvard-Smithsonian Center for Astrophysics, 60 Garden St., Cambridge, MA 02138, USA\\
$^{6}$Mullard Space Science Laboratory, University College London, Holmbury St Mary, Dorking, Surrey RH5 6NT, UK\\
$^{7}$Physik-Institut, University of Zurich, Winterthurerstrasse 190, 8057 Z\"urich, Switzerland}}

@@

\begin{abstract}
We measure the shape and alignment of the stellar and dark matter mass distribution in 11 strong-lens galaxies, finding \cb{over the range of several multiples of the half-light radii} that the dark matter haloes are rounder than the stellar mass distributions. Averaging over larger radii, the \cb{ellipticity, as defined using the ratio of the semi-minor and semi-major axes, increases both in their dark matter and stellar components.} While dark matter haloes are never more elliptical than $e_{dm} = 0.3$, their stars can extend to $e_* > 0.3$. Three systems, in particular, have a high stellar ellipticity ($e_* > 0.4$) and correspondingly high alignment between luminous and dark matter distribution. One of these -- {\it B1608+656} -- is a known merging pair; we suggest that the other two ({\it B0712+472} and {\it B2016+112}) may also be recent post-merger systems. Galaxies with high dark matter ellipticity and weak external shear show strong alignment between light and dark; those with strong shear ($\gamma \gtrsim 0.1$) can be highly misaligned. This is reassuring since isolated misaligned galaxies are expected to be unstable.

\cb{Our results provide a new constraint on galaxy formation models that must explain the origin of both very round dark matter haloes and \cb{misaligned} strong-lens systems.}

@@

The ellipticity and shape of the stellar component relative to their host dark matter halo encodes information both about our \cb{current cosmological model $\Lambda$CDM} and galaxy formation \citep[e.g.][]{1994ApJ...431..617D,2001ApJ...551..294I,2004ApJ...611L..73K,2007MNRAS.378...55M,2007arXiv0707.0737D,2012MNRAS.424L..16L,2014JPhG...41f3101R}. `Dark-matter-only' (DMO) simulations in $\Lambda$CDM predict dark matter haloes that are triaxial \citep{1991ApJ...378..496D,1992ApJ...399..405W,1996ApJ...462..563N,2002ApJ...574..538J}, with mean `shape parameter' $\langle q \rangle = (b+c)/2a \sim 0.8$ (where $a > b > c$ are the long, intermediate and short axes of the figure; \citealt{2007MNRAS.378...55M}). This corresponds to a typically {\it prolate} halo. However, including `baryons' (stars and gas) in the models produces haloes that are significantly rounder and -- at least for disc galaxies -- well-aligned with the light distribution \citep{1991ApJ...377..365K,1994ApJ...431..617D,2007arXiv0707.0737D}. Halo shapes and alignments also constrain alternative gravity models \citep{2001MNRAS.327..552M,2004ApJ...610L..97H,2005MNRAS.361..971R,2012PhRvD..86h3507F,2013MNRAS.434.2971D}. If stars dominate the mass of the galaxy, we expect the light and mass distribution to be highly correlated; if dark matter is present, however, such correlations can\cb{, at least in principle,} be broken.

@@

Weak lensing can also be used to probe the shape and alignment between the luminous and dark matter distributions, but only for `stacked' galaxies \citep{2000astro.ph..6281B,2000ApJ...538L.113N}. \citet{2004ApJ...606...67H} applied this idea to data from the Red-Sequence Cluster Survey (RCS) to measure the first weak lensing signal of halo flattening. They found that dark matter haloes appear to be rounder than the stellar mass distribution, with some weak evidence for alignment. Both measurements are challenging however, and more recent data appear to be at odds with this early work\cb{, at least for galaxies in general. They however are in agreement concerning `red' galaxies, which are predominantly early-type galaxies, also favouring dark matter haloes that are more elliptical than their stellar mass components \citep{2006MNRAS.370.1008M,2007ApJ...669...21P,2012A&A...545A..71V}.}

Recently \cite{2014arXiv1401.7990C} introduced a new non-parametric lens tool, \Glass. Applying this to a large suite of mock data, we showed that mass and light can only reliably be disentangled in lens systems if: i) there are at least four images; and ii) time delay data are available and/or the stellar mass contributes significantly to the potential. In this paper, we collate data of the above quality, compiling a sample of 11 strong lens galaxies. We apply \Glass\ to these lenses to non-parametrically measure the shape and alignment of the stars and {\it dark matter} in these lens galaxies, for the first time. This differs from previous works that have all compared the light distribution with the total mass, rather than the dark matter. Since the stellar component often dominates the central potential, the total mass naturally correlates with the light, potentially masking theoretically interesting results about the dark matter distribution. Our comparison between the stellar and dark matter components is made possible by the fact that \Glass\ uses the stellar mass distribution as a prior on the mass map, ensuring that the dark matter \cb{map} is always positive.

@@

\begin{enumerate}
\item A new uniform sampling algorithm for high dimensional spaces \citep{2012MNRAS.425.3077L}. This allows for large ensembles of $\gtrsim$10,000 models to be efficiently generated. 
\item With so many models in the final ensemble, we can afford to apply non-linear constraints (for example stellar kinematic data; or the removal of models with spurious extra images) to accept/reject models in a post-processing step.
\item The central region of the mass map can have a higher resolution to more efficiently capture steep models.
\item Stellar density can be used as an additional lower-bound constraint on the models. 
\end{enumerate}

@@

\begin{enumerate} 
\item Observational priors (redshifts, image positions relative to lens position).
\item Image parity is enforced.
\item A reconstructed stellar mass map as lower mass bounds in every pixel.
\item The local density gradient always points within 50$^{\circ}$ of the centre of the distribution, and the circularly-averaged density never increases outwards.
\item \cb{The fixed, flat cosmology with matter density $\Omega_{m} = 0.28$, the dark energy density $\Omega_{\Lambda} = 0.72$, and the inverse Hubble constant $H_{0}^{-1} = 13.7$ Gyr.}
\item A flat prior on the magnitude of the external shear $|\gamma|$ on $[0,1)$.
\item \cb{The radius of the lens is set to be 12 pixels.}
\end{enumerate}

@@

\begin{enumerate}
\item[(viii)] A flat prior on measured time delays within the $\pm 1\sigma$-interval.
\item[(ix)] Point symmetry $\kappa(x,y) = \kappa(-x,-y)$, where the surface mass profile can be assumed to be symmetric ((3+1)-quad configurations and the double \textit{Q0957}).

@@

\subsection{Measuring the shape and alignment of a lens}\label{sec:shapemethod}
We employ a shape definition related to the definition used by \cite{2014arXiv1401.7990C}. The ellipticity parameter $e$ is defined as
\begin{equation}\label{eq:shapeestimate}
    e \equiv 1 - \frac{\lambda_{2}}{\lambda_{1}},
\end{equation}
\cb{where $\lambda_{1}$ and $\lambda_{2}$ ($\lambda_{1} \geq \lambda_{2}$) are the eigenvalues of the inertia tensor $I$ of the mass distribution $M(\boldsymbol{\theta})$. It is defined as}
\begin{equation}\label{eq:inertiatensor}
    I = \begin{pmatrix} \sum_\theta M(\boldsymbol{\theta})\theta^{2}_{y} & -\sum_\theta M(\boldsymbol{\theta})\theta_{x}\theta_{y} \\
                        -\sum_\theta M(\boldsymbol{\theta})\theta_{x}\theta_{y} & \sum_\theta M(\boldsymbol{\theta})\theta^{2}_{x} \end{pmatrix}.
\end{equation}
We use these estimators to quantify the shape of the stellar mass distribution and the dark matter distribution. We estimate the latter by reconstructing the total mass distribution with a lower mass prior from the stellar mass distribution and then subtracting the latter off. The shapes of the distributions are computed for all the pixels up to the outermost lensed image and for the pixels within different multiples of \cb{$R_e$, defined as the half-light radius} measured with respect to the Petrosian flux. As terms in Eq.~\ref{eq:inertiatensor} are weighted with $r^2$, the outer pixels tend to dominate the estimate of the averaged shape within the considered radius. The shape estimates do not necessarily need to be the same within different radii -- e.g. effects such as isophotal twist \citep[e.g.][and references therein]{1978ComAp...8...27B} can cause differences in the shape estimates between the central and the outer parts of the system.

\cb{Throughout this paper, we generate in each case an ensemble of 10,000 mass models. Each of these models is analysed individually, yielding a distribution of constrained quantities.}

\section{Data}\label{sec:data}
Stellar mass maps are taken from \cite{2011ApJ...740...97L}. In that work, the surface brightness distribution of the galaxy images in all available HST bands was first fitted using {\sc Galfit v2.03b} \citep{2002AJ....124..266P}, and with the resulting flux maps stellar mass maps using the synthesis of the stellar-population (SPS) models of \cite{2003MNRAS.344.1000B} were obtained. The population synthesis is marginalised over the star formation epoch and time scales and over the stellar metallicity. The \cb{initial mass function (IMF)} was taken to be the log-normal form given by \cite{2003PASP..115..763C}; a more bottom-heavy IMF would change the mass normalisation \citep[cf.][]{2014ApJ...793...96S} but not the shape of the inferred stellar mass distribution \citep[unless the IMF presents significant intrinsic deviations locally, see e.g.][]{2015MNRAS.447.1033M}. We note that the effect of recent claims towards a non-universal IMF on stellar \cb{mass-to-light ratio (M/L)} is still under debate. \cb{While for massive elliptical galaxies, dynamical studies seem to point at contributions at the level of the Salpeter IMF or slightly heavier \citep{2013MNRAS.432.1862C}, spectral line strength constraints can still accommodate a wider range of the stellar M/L depending on the functional form of the IMF \citep{2013MNRAS.429L..15F}. For at least one lens in our sample, \textit{Q2237+030}, the lensing estimates on the mass distribution are inconsistent with a Salpeter IMF \citep{2010MNRAS.409L..30F}. We additionally tested varying the IMF for one lens ({\it B1422+231}), finding that Salpeter IMF is too heavy and gives no solutions. We will perform a more systematic study of the IMF and varying stellar M/L in a forthcoming publication.}
% also the lens system \textit{B1422+231} is incompatible with a Salpeter IMF. We will discuss constraints on different functional forms of the IMF and stellar M/L by strong-lens systems reconstructed non-parametrically with \Glass\ in a future publication.}

The original sample we consider, is presented in \cite{2011ApJ...740...97L} and consists of 21 lens systems. It is a subsample of the CfA-Arizona Space Telescope LEns Survey\footnote{http://www.cfa.harvard.edu/castles/} (CASTLES) sample. As shown in \cite{2014arXiv1401.7990C}, the recovery of the shape information of the lens is only good if there are at least four images. Therefore, we only consider the quad lenses and the twin double \textit{Q0957+561}. This \cb{yields a sample consisting} of 11 lenses. We discuss each of these systems, next.

@@

\textit{RXJ0911+0551} (hereafter \textit{0911}) is an almost spherical early-type galaxy \citep{1997A&A...317L..13B,2012A&A...538A..99S}, and has measured time delays \citep{2002ApJ...572L..11H}. \cb{It lies on the outskirts of a cluster of which X-ray emission can be detected yielding a temperature of 2.3 keV \citep{2001ApJ...555....1M}.} There is also a satellite galaxy to the northwest \citep{2000ApJ...544L..35K}.

@@

\textit{B1422+231} (hereafter \textit{1422}) is an early-type galaxy \citep{1992MNRAS.259P...1P,1994AJ....107...28Y}. Although time delays have been reported \citep{2001MNRAS.326.1403P}, it is possible that the measurements are aliases of the actual values \citep{2003AJ....126...29R} and hence are not included in our analysis. The lens is part of a group with 16 spectroscopically confirmed member galaxies \citep{2006ApJ...641..169M}. Using newer data, additional members could be \cb{identified} \citep{2011ApJ...726...84W}. \cite{2004ApJ...610..686G} also detect X-ray emission from the corresponding group at a temperature of 1.0 keV.

\textit{B1608+656} (hereafter \textit{1608}) was first reported by \cite{1995ApJ...447L...5M}. It consists of two merging galaxies. The main galaxy is an early-type galaxy which is disrupted by a smaller, probably late-type galaxy \citep{2003ApJ...584..100S}. The system has measured time delays \citep{2002ApJ...581..823F}. The environment has been analysed and a group with 8 (9 if merging galaxies are counted individually) was \cb{identified} \citep{2006ApJ...642...30F}. \cb{However, no} significant X-ray emission was detected from the surrounding group \citep{2005ApJ...625..633D}. The data seem to indicate four other groups along the line of sight.

\textit{MG2016+112} (hereafter \textit{2016}) is a giant elliptical galaxy \citep{1984Sci...223...46L,1986AJ.....91..991S}. It is the farthest lens we consider in this sample. It lies in a cluster that consists of 69 probable galaxies \citep{2003MNRAS.344..337T}. The cluster shows a high density of galaxies close to the lens in a south-east direction. Two lensed images are \cb{very close} in this fold lens, but could be resolved by \cite{2009MNRAS.394..174M}.

\textit{B2045+265} (hereafter \textit{2045}) is probably an elliptical galaxy \citep{2007MNRAS.378..109M}. It was initially classified as a late-type Sa galaxy \citep{1999AJ....117..658F}, however the velocity dispersion seems too high. As the source redshift is rather low, a large lens mass is required. \cite{2007MNRAS.378..109M} therefore conclude that it is more likely an elliptical galaxy. To the west of the lens, a group was found at a similar redshift as the lens \citep{1999AJ....117..658F}. There is furthermore evidence for a dwarf satellite galaxy \citep{2007MNRAS.378..109M}. As the measurements are \cb{however} inconclusive, we choose to not include this satellite in our models.

\textit{Q2237+030} (hereafter \textit{2237}) is a barred spiral \citep{1988AJ.....95.1331Y}. With a redshift of just $z\sim0.04$ it is the closest lens \cb{system} of the sample. Due to its low redshift, the probed physical scales are small, making the bulge component of the galaxy dominant. \cb{This makes the treatment of the stellar populations easier, as derivations of the stellar M/L are more difficult in the presence of star forming regions typical of disc galaxies, where there are no simple stellar populations usually assumed in SPS models.}

@@

\begin{table}
  \begin{center}
    \begin{tabular}{l r r r r l}
      Lens    & \multicolumn{1}{c}{$z_{L}$} & \multicolumn{1}{c}{$z_{S}$} & \multicolumn{1}{c}{$\Delta\theta$ [kpc]} & \multicolumn{1}{c}{$R_{L}$/$R_{e}$} & Environment \\ \hline
      0047 & 0.485 & 3.60 & 12.82 & $1.45\pm0.04$ & G(9) \\
      0414 & 0.960 & 2.64 & 16.01 & $1.85\pm0.05$ & ... \\
      0712 & 0.410 & 1.34 & 6.82  & $1.15\pm0.03$ & ... \\
      0911 & 0.769 & 2.8  & 23.16 & $3.09\pm0.05$ & C \\
      0957 & 0.356 & 1.41 & 29.98 & $3.51\pm0.04$ & C \\
      1115 & 0.310 & 1.72 & 10.76 & $2.86\pm0.06$ & G(13) \\
      1422 & 0.337 & 3.62 & 6.02  & $4.49\pm0.06$ & G(17) \\
      1608 & 0.630 & 1.39 & 13.92 & $1.82\pm0.01$ & G(8) \\
      2016 & 1.010 & 3.3  & 26.22 & $6.12\pm0.14$ & C(69) \\
      2045 & 0.870 & 1.28 & 14.46 & $1.48\pm0.03$ & ... \\
      2237 & 0.039 & 1.7  & 1.40  & $0.89\pm0.01$ & ... \\
    \end{tabular}
    \caption[width=\linewidth]{The most relevant lens properties for this work are listed \citep[for an expanded version of this table see][]{2011ApJ...740...97L}. References can be found in \S\ref{sec:data}. The columns are from left to right: the lens redshift $z_L$; the redshift of the source $z_S$; the maximum angular separation between two images $\Delta\theta$; the ratio of the radius of the outermost image $R_L$ and the effective Petrosian half-light radius of the lens $R_e$; and information on the environment of the lens. C and G denote a known cluster or group environment, respecitvely. The number in the brackets is the number of confirmed members. \cb{Ellipsis indicate lens galaxies with no group members identified so far.}}
    \label{tab:lensproperties}
  \end{center}
\end{table}

>>>>

\def\Glass{{\sc Glass}}
\def\PixeLens{{\sc PixeLens}}

@@

\author[Bruderer et al.]
{\parbox{\textwidth}{Claudio Bruderer,$^{1}$\thanks{E-mail: \texttt{claudio.bruderer@phys.ethz.ch}}
Justin I. Read,$^{1,2}$
Jonathan P. Coles,$^{3,4}$
Dominik Leier,$^{5}$
Emilio E. Falco,$^{6}$
Ignacio Ferreras$^{7}$ and
Prasenjit Saha$^{4,8}$}\vspace{0.4cm}\\
\parbox{\textwidth}{$^{1}$Institute for Astronomy, Department of Physics, ETH Zurich, Wolfgang-Pauli-Strasse 27, 8093 Z\"urich, Switzerland\\
$^{2}$Department of Physics, University of Surrey, Guildford, GU2 7XH, UK\\
$^{3}$Exascale Research Computing Lab, Campus Teratec, 2 Rue de la Piquetterie, 91680 Bruyeres-le-Chatel, France\\
$^{4}$Physik-Institut, University of Zurich, Winterthurerstrasse 190, 8057 Z\"urich, Switzerland\\
$^{5}$Dipartimento di Fisica e Astronomia, Universit\`{a} di Bologna, viale Berti Pichat 6/2, 40127, Bologna, Italy\\
$^{6}$Harvard-Smithsonian Center for Astrophysics, 60 Garden St., Cambridge, MA 02138, USA\\
$^{7}$Mullard Space Science Laboratory, University College London, Holmbury St Mary, Dorking, Surrey RH5 6NT, UK\\
$^{8}$Institute for Computational Science, University of Zurich, Winterthurerstrasse 190, 8057 Z\"urich, Switzerland}}

@@

\begin{abstract}
We measure the shape and alignment of the stellar and dark matter mass distribution in 11 strong-lens galaxies, finding over the range of several multiples of the half-light radii that the dark matter haloes are rounder than the stellar mass distributions. Averaging over larger radii, the ellipticity, as defined using the ratio of the semi-minor and semi-major axes, increases both in their dark matter and stellar components. While dark matter haloes are never more elliptical than $e_{dm} = 0.3$, their stars can extend to $e_* > 0.3$. Three systems, in particular, have a high stellar ellipticity ($e_* > 0.4$) and correspondingly a high alignment between the luminous and the dark matter distribution. One of these -- {\it B1608+656} -- is a known merging pair; we suggest that the other two ({\it B0712+472} and {\it B2016+112}) may also be recent post-merger systems. Galaxies with high dark matter ellipticity and weak external shear show strong alignment between light and dark; those with strong shear ($\gamma \gtrsim 0.1$) can be highly misaligned. This is reassuring since isolated misaligned galaxies are expected to be unstable.

Our results provide a new constraint on galaxy formation models that must explain the origin of both very round dark matter haloes and misaligned strong-lens systems.

@@

The ellipticity and shape of the stellar component relative to their host dark matter halo encodes information both about our current cosmological model $\Lambda$CDM and galaxy formation \citep[e.g.][]{1994ApJ...431..617D,2001ApJ...551..294I,2004ApJ...611L..73K,2007MNRAS.378...55M,2007arXiv0707.0737D,2012MNRAS.424L..16L,2014JPhG...41f3101R}. `Dark-matter-only' (DMO) simulations in $\Lambda$CDM predict dark matter haloes that are triaxial \citep{1991ApJ...378..496D,1992ApJ...399..405W,1996ApJ...462..563N,2002ApJ...574..538J}, with mean `shape parameter' $\langle q \rangle = (b+c)/2a \sim 0.8$ (where $a > b > c$ are the long, intermediate and short axes of the figure; \citealt{2007MNRAS.378...55M}). This corresponds to a typically {\it prolate} halo. However, including `baryons' (stars and gas) in the models produces haloes that are significantly rounder and -- at least for disc galaxies -- well-aligned with the light distribution \citep{1991ApJ...377..365K,1994ApJ...431..617D,2007arXiv0707.0737D}. Halo shapes and alignments also constrain alternative gravity models \citep{2001MNRAS.327..552M,2004ApJ...610L..97H,2005MNRAS.361..971R,2012PhRvD..86h3507F,2013MNRAS.434.2971D}. If stars dominate the mass of the galaxy, we expect the light and mass distribution to be highly correlated; if dark matter is present, however, such correlations can, at least in principle, be broken.

@@

Weak lensing can also be used to probe the shape and alignment between the luminous and dark matter distributions, but only for `stacked' galaxies \citep{2000astro.ph..6281B,2000ApJ...538L.113N}. \citet{2004ApJ...606...67H} applied this idea to data from the Red-Sequence Cluster Survey (RCS) to measure the first weak lensing signal of halo flattening. They found that dark matter haloes appear to be rounder than the stellar mass distribution, with some weak evidence for alignment. Both measurements are challenging however, and more recent data appear to be at odds with this early work, at least for galaxies in general. They however are in agreement concerning `red' galaxies, which are predominantly early-type galaxies, also favouring dark matter haloes that are more elliptical than their stellar mass components \citep{2006MNRAS.370.1008M,2007ApJ...669...21P,2012A&A...545A..71V}.

Recently, \cite{2014arXiv1401.7990C} introduced a new non-parametric lens tool, \Glass. Applying this to a large suite of mock data, we showed that mass and light can only reliably be disentangled in lens systems if: i) there are at least four images; and ii) time delay data are available and/or the stellar mass contributes significantly to the potential. In this paper, we collate data of the above quality, compiling a sample of 11 strong lens galaxies. We apply \Glass\ to these lenses to non-parametrically measure the shape and alignment of the stars and {\it dark matter} in these lens galaxies, for the first time. This differs from previous works that have all compared the light distribution with the total mass, rather than the dark matter. Since the stellar component often dominates the central potential, the total mass naturally correlates with the light, potentially masking theoretically interesting results about the dark matter distribution. Our comparison between the stellar and dark matter components is made possible by the fact that \Glass\ uses the stellar mass distribution as a prior on the mass map, ensuring that the dark matter map is always positive.

@@

\begin{enumerate}
\item A new uniform sampling algorithm for high dimensional spaces \citep{2012MNRAS.425.3077L} is employed. This allows for large ensembles of $\gtrsim$10,000 models to be efficiently generated. 
\item With so many models in the final ensemble, we can afford to apply non-linear constraints (for example stellar kinematic data; or the removal of models with spurious extra images) to accept/reject models in a post-processing step.
\item The central region of the mass map can have a higher resolution to more efficiently capture steep models.
\item Stellar density can be used as an additional lower-bound constraint on the models. 
\end{enumerate}

@@

\begin{enumerate} 
\item Observational priors (redshifts, image positions relative to lens position).
\item Image parity to be enforced.
\item A reconstructed stellar mass map that provides a lower mass bound to each pixel.
\item The local density gradient to be always pointing within 50$^{\circ}$ of the centre of the distribution, and the circularly-averaged density to never increase outwards.
\item The fixed, flat cosmology with matter density $\Omega_{m} = 0.28$, the dark energy density $\Omega_{\Lambda} = 0.72$, and the inverse Hubble constant $H_{0}^{-1} = 13.7$ Gyr.
\item A flat prior on the magnitude of the external shear $|\gamma|$ on $[0,1)$.
\item The radius of the lens to be set to 12 pixels.
\end{enumerate}

@@

\begin{enumerate}
\item[(viii)] A flat prior on measured time delays within the $\pm 1\sigma$-interval.
\item[(ix)] Point symmetry $\kappa(x,y) = \kappa(-x,-y)$, where the surface mass profile can be assumed to be symmetric ((3+1)-quad configurations and the double \textit{Q0957+561}).

@@

\subsection{Measuring the shape and alignment of a lens}\label{sec:shapemethod}
We employ a shape definition related to the definition used by \cite{2014arXiv1401.7990C}. The ellipticity parameter $e$ is defined as
\begin{equation}\label{eq:shapeestimate}
    e \equiv 1 - \frac{\lambda_{2}}{\lambda_{1}},
\end{equation}
where $\lambda_{1}$ and $\lambda_{2}$ ($\lambda_{1} \geq \lambda_{2}$) are the eigenvalues of the inertia tensor $I$ of the mass distribution $M(\boldsymbol{\theta})$. It is defined as
\begin{equation}\label{eq:inertiatensor}
    I = \begin{pmatrix} \sum_\theta M(\boldsymbol{\theta})\theta^{2}_{y} & -\sum_\theta M(\boldsymbol{\theta})\theta_{x}\theta_{y} \\
                        -\sum_\theta M(\boldsymbol{\theta})\theta_{x}\theta_{y} & \sum_\theta M(\boldsymbol{\theta})\theta^{2}_{x} \end{pmatrix}.
\end{equation}
We use these estimators to quantify the shape of the stellar mass distribution and the dark matter distribution. We estimate the latter by reconstructing the total mass distribution with a lower mass prior from the stellar mass distribution and then subtracting the latter off. The shapes of the distributions are computed for all the pixels up to the outermost lensed image and for the pixels within different multiples of $R_e$, defined as the half-light radius measured with respect to the Petrosian flux. As terms in Eq.~\ref{eq:inertiatensor} are weighted with $r^2$, the outer pixels tend to dominate the estimate of the averaged shape within the considered radius. The shape estimates do not necessarily need to be the same within different radii -- e.g. effects such as isophotal twist \citep[e.g.][and references therein]{1978ComAp...8...27B} can cause differences in the shape estimates between the central and the outer parts of the system.

Throughout this paper, we generate in each case an ensemble of 10,000 mass models. Each of these models is analysed individually, yielding a distribution of constrained quantities.

\section{Data}\label{sec:data}
Stellar mass maps are taken from \cite{2011ApJ...740...97L}. In that work, the surface brightness distribution of the galaxy images in all available HST bands was first fitted using {\sc Galfit v2.03b} \citep{2002AJ....124..266P}, and with the resulting flux maps stellar mass maps using the synthesis of the stellar-population (SPS) models of \cite{2003MNRAS.344.1000B} were obtained. The population synthesis is marginalised over the star formation epoch and time scales and over the stellar metallicity. The initial mass function (IMF) was taken to be the log-normal form given by \cite{2003PASP..115..763C}; a more bottom-heavy IMF would change the mass normalisation \citep[cf.][]{2014ApJ...793...96S} but not the shape of the inferred stellar mass distribution \citep[unless the IMF presents significant intrinsic deviations locally, see e.g.][]{2015MNRAS.447.1033M}. We note that the effect of recent claims towards a non-universal IMF on stellar mass-to-light ratio (M/L) is still under debate. While for massive elliptical galaxies, dynamical studies seem to point at contributions at the level of the Salpeter IMF or slightly heavier \citep{2013MNRAS.432.1862C}, spectral line strength constraints can still accommodate a wider range of the stellar M/L depending on the functional form of the IMF \citep{2013MNRAS.429L..15F}. For at least one lens in our sample, \textit{Q2237+030}, the lensing estimates on the mass distribution are inconsistent with a Salpeter IMF \citep{2010MNRAS.409L..30F}. We additionally tested varying the IMF for one lens ({\it B1422+231}), finding that Salpeter IMF is too heavy and gives no solutions. We will perform a more systematic study of the IMF and varying stellar M/L in a forthcoming publication.

The original sample we consider, is presented in \cite{2011ApJ...740...97L} and consists of 21 lens systems. It is a subsample of the CfA-Arizona Space Telescope LEns Survey\footnote{http://www.cfa.harvard.edu/castles/} (CASTLES) sample. As shown in \cite{2014arXiv1401.7990C}, the recovery of the shape information of the lens is only good if there are at least four images. Therefore, we only consider the quad lenses and the twin double \textit{Q0957+561}. This yields a sample consisting of 11 lenses. We discuss each of these systems, next.

@@

\textit{RXJ0911+0551} (hereafter \textit{0911}) is an almost spherical early-type galaxy \citep{1997A&A...317L..13B,2012A&A...538A..99S}, and has measured time delays \citep{2002ApJ...572L..11H}. It lies on the outskirts of a cluster of which X-ray emission can be detected yielding a temperature of 2.3 keV \citep{2001ApJ...555....1M}. There is also a satellite galaxy to the northwest \citep{2000ApJ...544L..35K}.

@@

\textit{B1422+231} (hereafter \textit{1422}) is an early-type galaxy \citep{1992MNRAS.259P...1P,1994AJ....107...28Y}. Although time delays have been reported \citep{2001MNRAS.326.1403P}, it is possible that the measurements are aliases of the actual values \citep{2003AJ....126...29R} and hence are not included in our analysis. The lens is part of a group with 16 spectroscopically confirmed member galaxies \citep{2006ApJ...641..169M}. Using newer data, additional members could be identified \citep{2011ApJ...726...84W}. \cite{2004ApJ...610..686G} also detect X-ray emission from the corresponding group at a temperature of 1.0 keV.

\textit{B1608+656} (hereafter \textit{1608}) was first reported by \cite{1995ApJ...447L...5M}. It consists of two merging galaxies. The main galaxy is an early-type galaxy which is disrupted by a smaller, probably late-type galaxy \citep{2003ApJ...584..100S}. The system has measured time delays \citep{2002ApJ...581..823F}. The environment has been analysed and a group with 8 (9 if merging galaxies are counted individually) was identified \citep{2006ApJ...642...30F}. However, no significant X-ray emission was detected from the surrounding group \citep{2005ApJ...625..633D}. The data seem to indicate four other groups along the line of sight.

\textit{MG2016+112} (hereafter \textit{2016}) is a giant elliptical galaxy \citep{1984Sci...223...46L,1986AJ.....91..991S}. It is the farthest lens we consider in this sample. It lies in a cluster that consists of 69 probable galaxies \citep{2003MNRAS.344..337T}. The cluster shows a high density of galaxies close to the lens in a south-east direction. Two lensed images are very close in this fold lens, but could be resolved by \cite{2009MNRAS.394..174M}.

\textit{B2045+265} (hereafter \textit{2045}) is probably an elliptical galaxy \citep{2007MNRAS.378..109M}. It was initially classified as a late-type Sa galaxy \citep{1999AJ....117..658F}, however the velocity dispersion seems too high. As the source redshift is rather low, a large lens mass is required. \cite{2007MNRAS.378..109M} therefore conclude that it is more likely an elliptical galaxy. To the west of the lens, a group was found at a similar redshift as the lens \citep{1999AJ....117..658F}. There is furthermore evidence for a dwarf satellite galaxy \citep{2007MNRAS.378..109M}. As the measurements are however inconclusive, we choose to not include this satellite in our models.

\textit{Q2237+030} (hereafter \textit{2237}) is a barred spiral \citep{1988AJ.....95.1331Y}. With a redshift of just $z\sim0.04$ it is the closest lens system of the sample. Due to its low redshift, the probed physical scales are small, making the bulge component of the galaxy dominant. This makes the treatment of the stellar populations easier, as derivations of the stellar M/L are more difficult in the presence of star forming regions typical of disc galaxies, where there are no simple stellar populations usually assumed in SPS models.

@@

\begin{table}
  \begin{center}
    \begin{tabular}{l r r r r l}
      Lens    & \multicolumn{1}{c}{$z_{L}$} & \multicolumn{1}{c}{$z_{S}$} & \multicolumn{1}{c}{$\Delta\theta$ [kpc]} & \multicolumn{1}{c}{$R_{L}$/$R_{e}$} & Environment \\ \hline
      0047 & 0.485 & 3.60 & 12.82 & $1.45\pm0.04$ & G(9) \\
      0414 & 0.960 & 2.64 & 16.01 & $1.85\pm0.05$ & ... \\
      0712 & 0.410 & 1.34 & 6.82  & $1.15\pm0.03$ & ... \\
      0911 & 0.769 & 2.8  & 23.16 & $3.09\pm0.05$ & C \\
      0957 & 0.356 & 1.41 & 29.98 & $3.51\pm0.04$ & C \\
      1115 & 0.310 & 1.72 & 10.76 & $2.86\pm0.06$ & G(13) \\
      1422 & 0.337 & 3.62 & 6.02  & $4.49\pm0.06$ & G(17) \\
      1608 & 0.630 & 1.39 & 13.92 & $1.82\pm0.01$ & G(8) \\
      2016 & 1.010 & 3.3  & 26.22 & $6.12\pm0.14$ & C(69) \\
      2045 & 0.870 & 1.28 & 14.46 & $1.48\pm0.03$ & ... \\
      2237 & 0.039 & 1.7  & 1.40  & $0.89\pm0.01$ & ... \\
    \end{tabular}
    \caption[width=\linewidth]{The most relevant lens properties for this work are listed \citep[for an expanded version of this table see][]{2011ApJ...740...97L}. References can be found in \S\ref{sec:data}. The columns are from left to right: the lens redshift $z_L$; the redshift of the source $z_S$; the maximum angular separation between two images $\Delta\theta$; the ratio of the radius of the outermost image $R_L$ and the effective Petrosian half-light radius of the lens $R_e$; and information on the environment of the lens. C and G denote a known cluster or group environment, respecitvely. The number in the brackets is the number of confirmed members. Ellipsis indicate lens galaxies with no group members identified so far.}
    \label{tab:lensproperties}
  \end{center}
\end{table}
